Title

### 2024-06-30 paper/main.tex
Updates

Inspired by this ranking approach in modification indices, in this paper we
develop a method to rank the CI tests using a conditional measure of
association. Our proposed approach uses the measure of association to find
correlated variables in the data that are not explained by the current model.
Using the measure, we can then rank the pair of variables between which adding
an edge would improve the explanation the most. Based on this information, the
user can select a potential pair of variables and based on their expert
knowledge decide the direction of the edge. Our main contributions in this
paper are as follows:
\begin{enumerate}
	\item We present a measure of (conditional) association for mixed data
		based on canonical correlations. This measure of association
		generalizes the widely used partial correlation to mixed data
		setting.
	\item Using this association measure, we present a procedure to suggest
		modifications (i.e., new edges, and deleting edges) to a given
		model such that the model can better explain the observed
		associations in the data. 
	\item Lastly, we present a web-browser based tool to allow researchers
		to apply this method to their own datasets.
\end{enumerate}

@@

@@

\section{Background and Related Work}
\label{sec:background}
We consider random variable $ X $ and $ Y $ with a possible multi-dimensional
variable $ \bm{Z} = \{ Z_1, \cdots, Z_k \} $. We consider these variables in
the mixed data setting where these variables can be any combination of
continuous, categorical, or ordinal unless specified. We use $ \rvert \bm{Z}
\rvert $ to represent the cardinality of the variable $ \bm{Z} $. We write $
\bm{x} = (x_1, \cdots, x_n) $ for a sample from $ X $ of size $ n $. We write
the expectation of a variable $ X $ as $ \mathbb{E}[X] $, conditional
expectations as $ \mathbb{E}[X | \bm{Z}] $.

@@

Using this mixed data partial association measure, we propose an iterative DAG
modification procedure. 
\todo[inline]{How similar is this to GES algorithm}
If the user desires to construct the model manually, they can also start with
an empty graph. Given a DAG $ G $ and a dataset $ D $, for each pair for
variable $ X $ and $ Y $ in $ G $ that does not have an edge between them, we
compute the partial association measure: $ \rho(X, Y; pa_G(X) \cup pa_G(Y)) $.
The measure of association in this case can be interpreted as the unexplained
observed association between them given the current model. We can then use this
to rank pair of variables which have very high unexplained association between
them. Based on the domain expertise, the user can then choose one of these
suggested pair of variables, decide the direction of the edge between the
variables, and add it to the model. Once the new edge is added we can recompute
the association of all other variables and $ X $ and $ Y $.

@@

\begin{theorem}
	If we have access to an oracle that can give correct edge directions, taking a
	greedy approach with the procedure above results in the correct graph
	being recovered.
\end{theorem}

@@

We analyzed the performance of this manual approach by comparing it with two
automated algorithms: PC and Hill-Climb Search. For the comparison we use
simulated data from a known DAG and compare how well the algorithm is able to
recover the original model using two metrics: Structural Hamming Distance (SHD)
and Structural Intervention Distance (SID). To simulate the data, we start with
a randomly generated DAG and use linear models with random effects to generate
the data. We repeat the data generation $ 30 $ times for each density value for
the DAG. As both PC and Hill-Climb Search can only give us a CPDAG, we compare
the orientation of these CPDAG which result in the best and worst case scores
on SHD and SID. This procedure is outlined in Algorithm 1.

@@

\begin{figure}[ht]
	\centering
	\includegraphics[scale=0.3]{../code/plots/llm_adult.png}
	\caption{\todo[inline]{Change figure to have consistent style}}
\end{figure}

>>>>

@@

Inspired by this ranking approach in modification indices, in this paper we
develop a method to rank the CI tests using a conditional measure of
association. Our proposed approach uses the measure of association to find
correlated variables in the data that are not explained by the current model.
Using the measure, we can then rank the pair of variables between which adding
an edge would improve the explanation the most. Based on this information, the
user can select a potential pair of variables and based on their expert
knowledge decide the direction of the edge. Our main contributions in this
paper are as follows:
\begin{enumerate}
	\item We present a measure of (conditional) association for mixed data
		based on canonical correlations. This measure of association
		generalizes the widely used partial correlation to mixed data
		setting.
	\item Using this partial association measure, we present a procedure to
		rank CI tests. This ranking can in turn then be used to
		prioritize addition or deletion of edges from the model.
	\item Lastly, we present a web-browser based tool to allow researchers
		to apply this method to their own datasets.
\end{enumerate}

@@

\section{Background and Related Work}
\label{sec:background}
We consider random variable $ X $ and $ Y $ with a possible multi-dimensional
variable $ \bm{Z} = \{ Z_1, \cdots, Z_k \} $. We consider these variables in
the mixed data setting where these variables can be any combination of
continuous, categorical, or ordinal unless specified. We also assume all
variables to be observed. We use $ \rvert \bm{Z} \rvert $ to represent the
cardinality of the variable $ \bm{Z} $. We write $ \bm{x} = (x_1, \cdots, x_n)
$ for a sample from $ X $ of size $ n $. We write the expectation of a variable
$ X $ as $ \mathbb{E}[X] $, conditional expectations as $ \mathbb{E}[X |
\bm{Z}] $.

@@

Using this mixed data partial association measure, we propose a method to rank
CI tests. If the user desires to construct the model manually, they can also
start with an empty graph, or start with a DAG outputted by a causal discovery
algorithm. Given a DAG $ G $ and a dataset $ D $, for each pair for variable
$ X $ and $ Y $ in $ G $ that does not have an edge between them, we compute
the partial association measure: $ \rho(X, Y; pa_G(X) \cup pa_G(Y)) $. The
measure of association in this case can be interpreted as the unexplained
observed association between them given the current model. We can then use this
to rank pair of variables which have very high unexplained association between
them. Based on the domain expertise, the user can then choose one of these
suggested pair of variables, decide the direction of the edge between the
variables, and add it to the model. Once the new edge is added we can recompute
the association of all other variables and $ X $ and $ Y $.

@@

\begin{theorem}
	If we have access to an oracle that can give correct edge directions, taking a
	greedy approach with the procedure above results in the correct graph
	being recovered. \todo[inline]{Reword and add proof}
\end{theorem}

@@

We analyzed the performance of this manual approach by comparing it with two
automated algorithms: PC and Greedy Equivalence Search. For the comparison we
use simulated data from a known DAG and compare how well the algorithm is able
to recover the original model using two metrics: Structural Hamming Distance
(SHD) and Structural Intervention Distance (SID). To simulate the data, we
start with a randomly generated DAG and use linear models with random effects
to generate the data. We repeat the data generation $ 30 $ times for each
density value for the DAG. As both PC and Hill-Climb Search can only give us a
CPDAG, we compare the orientation of these CPDAG which result in the best and
worst case scores on SHD and SID. This procedure is outlined in Algorithm 1.
best and worst case scores on SHD and SID. This procedure is outlined in
Algorithm 1.

@@

\begin{figure}
	\centering
	\begin{subfigure}{0.5\textwidth}
		\caption{\todo[inline]{Add the prompt used}}
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
		\centering
		\includegraphics[scale=0.3]{../code/plots/llm_adult.png}
		\caption{\todo[inline]{Change figure to have consistent style}}
	\end{subfigure}
\end{figure}
